---
title: "Heart Failure Inpatient Analysis"
author: "Harrison Nguyen"
date: "27/09/2021"
output:  
  bookdown::html_document2:
    code_folding: hide
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
    number_sections: true
bibliography: references.bib
link-citations: true
linkcolor: blue
vignette: >
  %\VignetteIndexEntry{Heart Failure Inpatient Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning=FALSE, message=FALSE
)
knitr::opts_knit$set(
  root.dir = here::here()
)
```

```{r import}

library(magrittr)
library(targets)
library(tidyverse)
library(kableExtra)
library(ggplot2)
library(ggpubr)
library(acs)
library(flextable)

library(hefd)
```

```{r}

proportion_bar_plot <- function(df,column,title){
  plot<- df %>%
     ggplot2::ggplot(., aes(x=!!rlang::sym(column),fill=!!rlang::sym(column))) +
    ggplot2::geom_bar(aes(y = ..count..),stat="count",alpha=0.5,show.legend = FALSE) +
    ggplot2::xlab(title) + 
    ggplot2::ylab("Count") + 
    ggtitle(paste0(title, " (n = ",nrow(df), ")")) +
    geom_text(
       aes(label=paste(round((..count..)/sum(..count..)*100,0),"%",sep=""),y= ..count..),
       stat='count',
       vjust=-0.25
   ) + 
  theme_minimal()
  
  return(plot)
}

conditional_bar_plot <- function(df,var1,var2,title){
  
  var1<- rlang::sym(var1)
  var2<- rlang::sym(var2)
  plot <- df%>%
  dplyr::count(!!var1,!!var2) %>%
  dplyr::group_by(!!var1) %>%
  dplyr::mutate(prop = prop.table(n)*100) %>%
  ggplot2::ggplot(., aes(x=!!var1,y=n,fill=!!var2)) + 
  geom_bar(stat="identity",position=position_dodge(0.7),width=0.7,alpha=0.5) +
  geom_text(aes(label= paste(round(prop,0),"%",sep=""),y=n),position=position_dodge(0.7),vjust=-0.5) +
  labs(y="Count",x=var1, title = paste0(title, " (N = ",nrow(df), ")")) +
  scale_fill_discrete(name = var2) + 
  theme_minimal()
  
  return(plot)
}

plot_histogram <- function(df,column, title,threshold = NULL){
  var1<- rlang::sym(column)

  df %<>% dplyr::filter(!is.na(!!var1))
  
  
  plot <- df %>%
  ggplot2::ggplot(., aes(x=!!var1,fill=!!var1)) + 
   geom_histogram(fill="lightblue",alpha=0.5) +
  labs(y="Count",x=var1, title = paste0(title, " (N = ",nrow(df), ")"))+
    scale_x_continuous(breaks = scales::pretty_breaks(n = 10))+
    stat_bin(aes(y=..count.., label=ifelse(..count..==0,"",..count..)), geom="text", vjust=-.5) +
  theme_minimal()
    

  
  if(!is.null(threshold)){
    plot <- plot + geom_vline(aes(xintercept=threshold),
            color="red", linetype="dashed", size=1)
  }
  
  return(plot)
}

plot_histogram_group <- function(df,var1,var2,title){
  var1<- rlang::sym(var1)
  var2<- rlang::sym(var2)

  df %<>% dplyr::filter(!is.na(!!var1)) %>%
    dplyr::filter(!is.na(!!var2))
  
  
  plot <- df %>%
  ggplot2::ggplot(., aes(x=!!var1,fill=!!var2,color=!!var2)) + 
   geom_histogram(aes(y=..density..),alpha=0.5, position="identity") +
    geom_density(alpha=0.3)+
  labs(y="Count",x=var1, title = paste0(title, " (N = ",nrow(df), ")"))+
    scale_x_continuous(breaks = scales::pretty_breaks(n = 10))+
  theme_minimal()
    
  
  return(plot)
}

plot_logical_columns <- function(df,var1,title){
  var1<- rlang::sym(var1)
  
  plot_data <- df %>% 
    dplyr::summarise_if(is.logical,sum) %>%
   tidyr::pivot_longer(everything(),values_to = "N",names_to = rlang::as_string(var1)) %>%
  dplyr::mutate(prop = N/nrow(df)*100)

  plot <- plot_data %>%
       ggplot2::ggplot(., aes(x=!!var1,y=N,fill=!!var1)) +
      ggplot2::geom_bar(position="dodge", stat="identity",show.legend = FALSE) +
    geom_text(aes(label= paste(round(prop,0),"%",sep=""),y=N),vjust=-0.5) + 
    labs(y="Count",x=var1, title = paste0(title, " (N = ",nrow(df), ")"))
  
  return(plot)
}

```

```{r}
plot_pie_chart <- function(df,var1,title){
  var1<- rlang::sym(var1)
  df %<>% 
    filter(!is.na(!!var1)) %>%
    group_by(!!var1) %>% # Variable to be transformed
    count() %>% 
    ungroup() %>% 
    mutate(perc = `n` / sum(`n`)) %>% 
    arrange(perc) %>%
    mutate(labels = scales::percent(perc))
  
  plot <- ggplot(df, aes(x = "", y = n, fill = !!var1)) +
    geom_col() +
    geom_text(aes(label = labels),
              position = position_stack(vjust = 0.5)) +
    coord_polar(theta = "y")
    labs(y="Count",x=var1, title = paste0(title, " (N = ",nrow(df), ")"))
    
  return(plot)
}

```

```{r data-load}
encounter_ref <- load_data("encounter_reference")
encounter <- load_data("encounter")
diagnosis <- load_data("diagnosis")
diagnosis_history <- load_data("diagnosis_history")
problem <- load_data("problem")
orders <- load_data("orders")
patient <- load_data("patient")
```


```{r}
query <- glue::glue("SELECT * from MEDICATION_ORDERS as x \\
            WHERE ENCNTR_ID in (SELECT PX9_ENCNTR_ID from ENCOUNTER_REFERENCE \\
             WHERE Q4_DIAGNOSIS = 1 and PROCESS_FLAG = 1 and HISTORY_FLAG = 0) and (ORIG_ORD_AS_FLAG = 1 OR ORIG_ORD_AS_FLAG=2)")
medication_order <- execute_query(query)
```

```{r}
location_string <- "Royal North Shore"
hf_diag <- diagnosis %>%
  dplyr::filter(stringr::str_detect(SOURCE_IDENTIFIER,"(?i)I50"))

qual_enc <- encounter %>%
  dplyr::filter(ENCNTR_ID %in% hf_diag$ENCNTR_ID & LOC_FACILITY_CD == location_string)

cohort <- qual_enc %>%
  dplyr::distinct(PERSON_ID)
```

# Introduction

This document will investigate encounters that have been coded with a heart failure related ICD10 code (I150) at **`r location_string`** Hospital with discharge dates between **`r as.Date(min(qual_enc$DISCH_DT_TM,na.rm = TRUE))`** to **`r as.Date(max(qual_enc$DISCH_DT_TM,na.rm = TRUE))`**. Thus, in a 2.5 year period, there are **`r nrow(qual_enc)`** encounters with unique **`r nrow(cohort)`** patients. Figure \@ref(fig:dist-enc-plot) shows the distribution of the number of encounters for each patient. 

The analysis of this cohort will be used to examine: 

1. Demographics and characteristics

* Total admissions by facility
* Age and sex distributions
* Numbers of patient journeys and transfers across the District
* Mode of admission and mode of separation
* Number of patients with Advanced Care Directive
* Previous hospitalisation in the past 3 years (for HF or other?, primary or additional diagnoses?)

2. Investigations

* Number of patients with at least 1 echocardiogram performed
* BNP and Iron profiles

3. Discharge management
* Number of patients with discharge letters
* Number of patients on relevant discharge medications or none
* Number of patients that are referred to the HF service

This feasibility analyses will be initially be performed for encounters at RNSH, then Ryde and Hornsby, and then patients that transfer between NSLHD facilities.

```{r dist-enc-plot,fig.cap="Distribution of the number of encounters for each patient."}
counts <- qual_enc %>%
  dplyr::group_by(PERSON_ID) %>%
  dplyr::summarise(N = dplyr::n())

proportion_bar_plot(counts,"N", "Number of encounters per patient")
```

```{r}
extract_hf_diagnosis <- function(df){
  df %<>%
  dplyr::group_by(PERSON_ID) %>%
  dplyr::summarise(
    `Congestive heart failure` = any(SOURCE_IDENTIFIER == "I50.0" | SOURCE_IDENTIFIER == "I50"),
    `Heart failure, unspecified` = any(SOURCE_IDENTIFIER == "I50.9"),
    `Left ventricular failure` = any(SOURCE_IDENTIFIER == "I50.1")
  ) %>%
  replace(is.na(.), FALSE)
  
  return(df)
}
```

## Cohort Stratification - Primary & Secondary Heart Failure Diagnosis

```{r}

hf_diag_primary <- hf_diag %>%
  dplyr::filter(DIAG_PRIORITY == 1 & ENCNTR_ID %in% qual_enc$ENCNTR_ID)

primary_temp <- cohort %>% 
  dplyr::left_join(hf_diag_primary,by="PERSON_ID") %>%
  dplyr::group_by(PERSON_ID) %>%
  dplyr::summarise(
    `Congestive heart failure` = any(SOURCE_IDENTIFIER == "I50.0" | SOURCE_IDENTIFIER == "I50"),
    `Heart failure, unspecified` = any(SOURCE_IDENTIFIER == "I50.9"),
    `Left ventricular failure` = any(SOURCE_IDENTIFIER == "I50.1")
  ) %>%
  replace(is.na(.), FALSE) %>%
  dplyr::mutate(`Only secondary HF diagnosis` = !`Congestive heart failure` & !`Heart failure, unspecified` & !`Left ventricular failure`)

hf_diag_secondary <- hf_diag %>%
  dplyr::filter(DIAG_PRIORITY != 1 & ENCNTR_ID %in% qual_enc$ENCNTR_ID)

secondary_temp <- cohort %>% 
  dplyr::left_join(hf_diag_secondary,by="PERSON_ID") %>%
  dplyr::group_by(PERSON_ID) %>%
  dplyr::summarise(
    `Congestive heart failure` = any(SOURCE_IDENTIFIER == "I50.0" | SOURCE_IDENTIFIER == "I50"),
    `Heart failure, unspecified` = any(SOURCE_IDENTIFIER == "I50.9"),
    `Left ventricular failure` = any(SOURCE_IDENTIFIER == "I50.1")
  ) %>%
  replace(is.na(.), FALSE) %>%
  dplyr::mutate(`Only primary HF diagnosis` = !`Congestive heart failure` & !`Heart failure, unspecified` & !`Left ventricular failure`)


hf_diag_primary_count <- hf_diag_primary %>%
  dplyr::group_by(ENCNTR_ID) %>%
  dplyr::summarise(N = dplyr::n()) %>%
  dplyr::filter(N > 1)
```  

Figure \@ref(fig:icd10-dist-plot) shows the distribution of primary and secondary heart failure ICD10 codes. In some cases, there are encounters with multiple heart failure primary diagnosis, (n = `r nrow(hf_diag_primary_count)`). An encounter can also be associated with both a primary and secondary heart failure ICD10. For example, in Figure \@ref(fig:icd10-dist-plot)**A** shows that **64**\% of encounters ONLY have a secondary diagnosis. Similarly  Figure \@ref(fig:icd10-dist-plot)**B** shows that **23**\% of encounters ONLY have a primary diagnosis. Therefore, **13**\% of encounters have both a  primary and secondary heart failure ICD10 diagnosis. 

For the given diagnosis, **congestive heart failure** is most prevalent, followed by **left ventricular failure** then **heart failure, unspecified**.
 

```{r icd10-dist-plot, fig.cap="Distribution of primary and secondary ICD10 diagnosis." }

p1 <- plot_logical_columns(primary_temp,"ICD10","Primary HF ICD10") + 
  scale_x_discrete(guide = guide_axis(n.dodge=2)) + 
  ggplot2::ylim(0,2000)

p2 <- plot_logical_columns(secondary_temp,"ICD10","Secondary HF ICD10") + 
  scale_x_discrete(guide = guide_axis(n.dodge=2))+ 
  ggplot2::ylim(0,2000)

ggpubr::ggarrange(p1,p2,labels= c("A","B"), ncol = 1, nrow = 2)
```
```{r}
n_only_prim <-  sum(secondary_temp$`Only primary HF diagnosis`)
```

Table \ref(tab: primary-diag-secondary-hf-tab) shows the primary diagnosis of those encounters with a secondary heart failure diagnosis (n = **`r nrow(secondary_temp) - n_only_prim`**). The data shows that these encounters are also commonly associated with **pneumonia**, as well as some other heart related issue such as **atrial fibrillation** or a **myocardial infarction**.

```{r primary-diag-secondary-hf-tab}
only_secondary_cohort <- secondary_temp %>% dplyr::filter(!`Only primary HF diagnosis`)

primary_diag_of_secondary <- diagnosis %>%
  dplyr::filter(!stringr::str_detect(SOURCE_IDENTIFIER,"(?i)I50") & 
                  PERSON_ID %in% only_secondary_cohort$PERSON_ID &
                  DIAG_PRIORITY == 1 &
                  SOURCE_VOCABULARY_CD == "ICD10-AM" &
                  ENCNTR_ID %in% qual_enc$ENCNTR_ID) 

primary_diag_of_secondary %>%
  dplyr::group_by(SOURCE_STRING) %>%
  dplyr::summarise(N = dplyr::n_distinct(PERSON_ID)) %>% 
  dplyr::mutate(prop = round(N/dplyr::n_distinct(primary_diag_of_secondary$PERSON_ID)*100,1)) %>%
  dplyr::arrange(desc(N)) %>% 
  dplyr::top_n(15) %>%
  knitr::kable(caption = "Top 15 primary diagnosis for those with only a secondary heart failure diagnosis.")

```

```{r}
primary_cohort <- cohort %>%
  dplyr::filter(PERSON_ID %in% hf_diag_primary$PERSON_ID) %>%
  dplyr::mutate(HAS_PRIMARY_HF = TRUE)


AGE_breaks <- c(-Inf,18, 25, 45, 65, 80, Inf)
  AGE_labels <- c("<18", "19-24", "25-44", "45-64", "65-80", "80+")
  
primary_enc <- encounter %>%
  dplyr::filter(ENCNTR_ID %in% hf_diag_primary$ENCNTR_ID) %>%
  dplyr::left_join(patient, by = "PERSON_ID") %>%
  dplyr::mutate(AGE = as.integer(difftime(ACTIVE_STATUS_DT_TM,BIRTH_DT_TM, units="weeks")/52.25)) %>%
  dplyr::mutate(AGE_BINNED = cut(AGE, breaks = AGE_breaks, labels = AGE_labels,
                                   right = FALSE)) %>%
    dplyr::mutate(LOS = difftime(DISCH_DT_TM,ARRIVE_DT_TM,units="days"))
```

The cohort that will be analysed for the rest of this document are those encounters with a primary heart failure diagnosis. 

# Demographics

With these selection and exclusion criteria, the cohort for this analysis will include  **`r nrow(hf_diag_primary)`** number of encounters for **`r nrow(primary_cohort)`** unique patients with the following demographics.

```{r}
total<- bind_rows(primary_enc, mutate(primary_enc, SEX_CD = "Total")) %>%
  group_by(SEX_CD) %>%
  summarise(
    N = n(),
    Age = round(mean(AGE)),
  ) %>%
  pivot_wider(names_from = "SEX_CD", values_from = c("N", "Age")) %>%
  mutate(p_Female = round(N_Female/N_Total*100, 1),
         p_Male = round(N_Male/N_Total*100, 1)) %>%
  select(
         Females = N_Female, p_Female,
         Males = N_Male, p_Male,
         N_Total,
         Age_Female,
         Age_Male,
         Age_Total) %>%
  flextable() %>%
  set_header_labels(Group = "", p_Female = "%", p_Male = "%", N_Total = "n",
                    Age_Female = "Age (F)", Age_Male = "Age (M)", Age_Total = "Mean Age")
total
```

Figure \@ref(fig:age-sex-plot) shows the distribution of sex according to age groups.

```{r age-sex-plot, fig.cap="Distribution of Sex conditioned on Age."}
conditional_bar_plot(primary_enc,"AGE_BINNED","SEX_CD", "Age and Sex") +
  scale_fill_discrete(name = "SEX") + 
  xlab("AGE")
```

Figure \@ref(fig:los-dis-plot) shows the distribution of the length of stay (LOS) marked with the average LOS (**`r round(as.numeric(mean(primary_enc$LOS,na.rm=TRUE)),1)`** days). Encounters with either a missing admission or discharge date have been excluded.

```{r los-dis-plot, fig.cap="Distribution of Length of Stay"}

plot_histogram(primary_enc,"LOS","Distribution of LOS",mean(primary_enc$LOS,na.rm=TRUE)) + xlab("LOS (days)") + xlim(0,50)
```
Figure \@ref(fig:encntr-type-plot) shows the distribution of encounters types. **All** primary heart failure coded encounters are **inpatient** encounters.

```{r encntr-type-plot, fig.cap="Distribution of Encounter Type."}
proportion_bar_plot(primary_enc,"ENCNTR_TYPE_CD","Distribution of Encounter Type.")
```
# Advanced Care Directive


```{r acd-tab}
known_to_hf <- problem %>%
  dplyr::filter(stringr::str_detect(SOURCE_STRING,"(?i)Known to heart failure")) %>%
  dplyr::select(PERSON_ID) %>%
  dplyr::mutate(HAS_ALERT = TRUE)

kk<- primary_cohort %>%
  merge(.,known_to_hf , by="PERSON_ID",all=TRUE) %>%
  replace(is.na(.), FALSE)

prop_tab <- prop.table(table(kk$HAS_PRIMARY_HF,kk$HAS_ALERT))
prop_tab %>% 
  knitr::kable(caption = "Contigency table of Primary HF ICD10 and Advanced Care directive (ACD).",digits=2) %>%
  kableExtra::add_header_above(c("Has Primary\n ICD10"=1,"Has ACD"=2))%>%
  kable_styling(position = "center",full_width = FALSE)

```
Table \@ref(tab:acd-tab) shows a contingency table of the number of patients with a primary ICD10 code who also have an Advanced Care Directive. **`r round(prop_tab[2,1]/sum(prop_tab[2,])*100,1)`**\% of patients with a primary HF ICD10 have an advanced care directive. Table xxx shows the cohort with either a primary or secondary HF diagnosis.

# Historical Diagnosis

Figure \@ref(fig:his-diag-plot) shows whether or not, each person in the cohort has had a an encounter with a HF ICD10 diagnosis within the last three years. 

```{r his-diag-plot, fig.cap="Distribution of historical diagnosis."}
his_hf_diag <- diagnosis_history %>%
  dplyr::filter(stringr::str_detect(SOURCE_IDENTIFIER,"(?i)I50") &
                  PERSON_ID %in% primary_cohort$PERSON_ID)

summary_his_diag <- extract_hf_diagnosis(his_hf_diag) %>%
  dplyr::right_join(primary_cohort,by="PERSON_ID") %>%
  dplyr::select(-HAS_PRIMARY_HF) %>%
  replace(is.na(.),FALSE) %>%
    dplyr::mutate(`No His. HF ICD10` = !`Congestive heart failure` & !`Heart failure, unspecified`
                  & !`Left ventricular failure`)
plot_logical_columns(summary_his_diag,"ICD10","Distribution of Historical ICD10")
```
# Orders

```{r}

echos <- orders %>%
  dplyr::filter(stringr::str_detect(CATALOG_CD,"(?i)cardio - echo") & ENCNTR_ID %in% primary_enc$ENCNTR_ID) %>%
  dplyr::group_by(ENCNTR_ID) %>%
  dplyr::arrange(desc(STATUS_DT_TM)) %>%
  dplyr::slice(1) %>%
  dplyr::select(ENCNTR_ID,CATALOG_CD,STATUS_DT_TM) %>%
  dplyr::right_join(
    dplyr::select(primary_enc,ENCNTR_ID,ARRIVE_DT_TM),by="ENCNTR_ID") %>%
  dplyr::mutate(HAS_ECHO = !is.na(CATALOG_CD),ECHO_DELTA = difftime(STATUS_DT_TM,ARRIVE_DT_TM,units = "days")) %>%
  dplyr::ungroup() %>%
  dplyr::select(-CATALOG_CD) %>%
  dplyr::rename(ECHO_START_DT_TM = STATUS_DT_TM)

```

```{r}

extract_path_result <- function(df,string,prefix,breaks=NULL,labels=NULL){
  df %<>% dplyr::filter(EVENT_CD == string) %>%
  dplyr::select(ENCNTR_ID,EVENT_START_DT_TM,RESULT_VAL,RESULT_UNITS_CD) %>%
    dplyr::mutate(RESULT_VAL = as.numeric(stringr::str_remove_all(RESULT_VAL, '\"')) )  %>% 
  dplyr::rename("{prefix}_EVENT_START_DT_TM" := EVENT_START_DT_TM,
                
                "{prefix}_RESULT_UNITS_CD" := RESULT_UNITS_CD)
  
  if(!is.null(breaks)){
    df %<>% dplyr::mutate("{prefix}_BIN" := cut(RESULT_VAL,breaks=breaks,labels=labels))
  }
  df %<>% dplyr::rename("{prefix}_RESULT_VAL" := RESULT_VAL)
  
  return(df)
}
```

```{r}
path_tests <- execute_query("SELECT * from PATHOLOGY_EVENT where CATALOG_CD = 'Brain Natriuretic Peptide' or CATALOG_CD = 'Iron Level'")

path_tests %<>%
  dplyr::filter(ENCNTR_ID %in% primary_enc$ENCNTR_ID) %>%
  dplyr::group_by(ENCNTR_ID,EVENT_CD) %>%
  dplyr::arrange(desc(EVENT_START_DT_TM)) %>%
  dplyr::slice(1) %>%
  dplyr::ungroup()



```
```{r order-prop-plot, fig.cap="Number and proportion of Encounters with a relevant heart failure order, echocardiogram, brain natriuretic peptide test, serum iron test and transferrin saturation test."}
hf_order<-echos  %>%
  dplyr::left_join(
    extract_path_result(path_tests,"Brain Natriuretic peptide (BNP )", "BNP", c(-Inf,450,900,Inf),c("< 450","450-900","900+")), by="ENCNTR_ID"
  ) %>%
  dplyr::left_join(
    extract_path_result(path_tests,"Iron", "IRON",c(-Inf,9,30.4,Inf),c("< 9","9-30.4","30.4+")), by="ENCNTR_ID"
  ) %>%
  dplyr::left_join(
    extract_path_result(path_tests,"Transferrin Saturation", "TRANSFERRIN_SAT",c(-Inf,20,Inf),c("< 20","20+")), by="ENCNTR_ID"
  ) %>%
  dplyr::mutate(HAS_BNP = !is.na(BNP_EVENT_START_DT_TM),
                HAS_IRON = !is.na(IRON_EVENT_START_DT_TM),
                HAS_TRANSFERRIN_SAT = !is.na(TRANSFERRIN_SAT_EVENT_START_DT_TM))

plot_logical_columns(hf_order,"ORDERS","Proportion of Encounter with Order")


```


```{r echo-time-plot,fig.cap="Distribution of the time taken for an echocardiogram."}

mean_delta <- mean(hf_order$ECHO_DELTA,na.rm = TRUE)
plot_histogram(hf_order %>% dplyr::filter(ECHO_DELTA < 75),
               "ECHO_DELTA","Distribution of time taken for Echo",
               mean_delta) + 
    xlab("Time take for echo since admission (days)")
```

Figure \@ref(fig:order-prop-plot) shows the proportion of encounters with a relevant heart failure order being echocardiogram, 
brain natriuretic peptide test, serum iron test and transferrin saturation test. Furthermore Figure \@ref(fig:echo-time-plot) shows the distribution of the time taken for an echocardiogram with a mean time of **`r round(mean_delta,1)`** days.


```{r path-test-plot,fig.cap="Results of pathology tests."}


p1<-plot_pie_chart(hf_order,"BNP_BIN","BNP Results by Group")
p2<-plot_pie_chart(hf_order,"IRON_BIN","BNP Results by Group")
p3<-plot_pie_chart(hf_order,"TRANSFERRIN_SAT_BIN","BNP Results by Group")

#ggpubr::ggarrange(p1,p2,p3,nrow=3,ncol=1,labels=c("A","B","C"))


p1
p2
p3
```

# Forms

```{r}
extract_form <- function(df,string,prefix){
  df %<>%
  dplyr::filter(DESCRIPTION == string) %>%
  dplyr::mutate(HAS_FORM = TRUE) %>%
  dplyr::group_by(PERSON_ID) %>%
  dplyr::arrange(desc(FORM_DT_TM)) %>%
  dplyr::slice(1) %>%
  dplyr::select(PERSON_ID,ENCNTR_ID,HAS_FORM,PARENT_ENTITY_ID,FORM_DT_TM) %>%
    dplyr::rename_with(.fn = ~ paste0(prefix,"_",.x))
  
  return(df)
}
  
```

```{r form-plot,fig.cap = "Proportion of Patients with an enrolment or referral form."}

hf_form <- execute_query("SELECT * from DCP_FORMS_ACTIVITY where DESCRIPTION = 'Heart Failure (MACARF) Referral Form' or DESCRIPTION = 'Heart Failure - Management of Cardiac Function Enrolment'")


form_primary_cohort <- primary_cohort %>%
  dplyr::left_join(
    extract_form(hf_form,'Heart Failure - Management of Cardiac Function Enrolment',"ENROLMENT"), by = c("PERSON_ID" = "ENROLMENT_PERSON_ID")
  ) %>%
  dplyr::left_join(
    extract_form(hf_form,'Heart Failure (MACARF) Referral Form',"REFERRAL"), by = c("PERSON_ID" = "REFERRAL_PERSON_ID")
  ) %>%
  tidyr::replace_na(list(ENROLMENT_HAS_FORM = FALSE, REFERRAL_HAS_FORM = FALSE))

plot_logical_columns(form_primary_cohort %>% dplyr::select(-HAS_PRIMARY_HF),"Form","Proportion with HF Form")
```

Figure \@ref(fig:form-plot) shows the proportion of patients who have had a referral form or enrolment form completed in the EMR. 

# Medication

```{r}
primary_meds <- medication_order %>%
  dplyr::filter(ENCNTR_ID %in% primary_enc$ENCNTR_ID)
primary_enc_meds <- primary_enc %>%
  dplyr::mutate(HAS_DISCH_MED = ENCNTR_ID %in% primary_meds$ENCNTR_ID)

```

Figure \@ref(fig:med-type-plot)**A** show the proportion of medications that were either a home or one that was ordered for discharged. Figure \@ref(fig:med-type-plot)**B** shows the proportion of the encounters that were given a discharge medication. 

```{r med-type-plot, fig.cap="The type of medication whether home or discharge medication."}
p1 <- proportion_bar_plot(primary_meds,"ORIG_ORD_AS_FLAG","Type of discharge medication\n")
p2 <- proportion_bar_plot(primary_enc_meds,"HAS_DISCH_MED","Encounters with Discharge Medications\n")

ggpubr::ggarrange(p1,p2,ncol=2,nrow=1,labels=c("A","B"))
```

Table \@ref(tab:med-class-tab) shows the top 15 medication classes given at discharge or home.

```{r med-class-tab}
primary_meds %>%
  dplyr::group_by(MEDICATION_CLASS) %>%
  dplyr::summarise(N = dplyr::n_distinct(PERSON_ID)) %>% 
  dplyr::mutate(prop = round(N/dplyr::n_distinct(primary_meds$PERSON_ID)*100,1)) %>%
  dplyr::arrange(desc(N)) %>% 
  dplyr::top_n(15) %>%
  knitr::kable(caption = "Top 15 discharge medication class.")
```
# Appendix

## Inclusion of Primary and Secondayr HF Diagnosis

```{r}
primary_cohort <- cohort %>%
  dplyr::filter(PERSON_ID %in% qual_enc$PERSON_ID) %>%
  dplyr::mutate(HAS_PRIMARY_HF = TRUE)


AGE_breaks <- c(-Inf,18, 25, 45, 65, 80, Inf)
  AGE_labels <- c("<18", "19-24", "25-44", "45-64", "65-80", "80+")
  
primary_enc <- qual_enc %>%
  dplyr::left_join(patient, by = "PERSON_ID") %>%
  dplyr::mutate(AGE = as.integer(difftime(ACTIVE_STATUS_DT_TM,BIRTH_DT_TM, units="weeks")/52.25)) %>%
  dplyr::mutate(AGE_BINNED = cut(AGE, breaks = AGE_breaks, labels = AGE_labels,
                                   right = FALSE)) %>%
    dplyr::mutate(LOS = difftime(DISCH_DT_TM,ARRIVE_DT_TM,units="days"))
```


```{r}
total<- bind_rows(primary_enc, mutate(primary_enc, SEX_CD = "Total")) %>%
  group_by(SEX_CD) %>%
  summarise(
    N = n(),
    Age = round(mean(AGE)),
  ) %>%
  pivot_wider(names_from = "SEX_CD", values_from = c("N", "Age")) %>%
  mutate(p_Female = round(N_Female/N_Total*100, 1),
         p_Male = round(N_Male/N_Total*100, 1)) %>%
  select(
         Females = N_Female, p_Female,
         Males = N_Male, p_Male,
         N_Total,
         Age_Female,
         Age_Male,
         Age_Total) %>%
  flextable() %>%
  set_header_labels(Group = "", p_Female = "%", p_Male = "%", N_Total = "n",
                    Age_Female = "Age (F)", Age_Male = "Age (M)", Age_Total = "Mean Age")
total
```

Figure \@ref(fig:age-sex-plot-sec) shows the distribution of sex according to age groups.

```{r age-sex-plot-sec, fig.cap="Distribution of Sex conditioned on Age."}
conditional_bar_plot(primary_enc,"AGE_BINNED","SEX_CD", "Age and Sex") +
  scale_fill_discrete(name = "SEX") + 
  xlab("AGE")
```

Figure \@ref(fig:los-dis-plot-sec) shows the distribution of the length of stay (LOS) marked with the average LOS (**`r round(as.numeric(mean(primary_enc$LOS,na.rm=TRUE)),1)`** days). Encounters with either a missing admission or discharge date have been excluded.

```{r los-dis-plot-sec, fig.cap="Distribution of Length of Stay"}

plot_histogram(primary_enc,"LOS","Distribution of LOS",mean(primary_enc$LOS,na.rm=TRUE)) + xlab("LOS (days)") + xlim(0,50)
```
Figure \@ref(fig:encntr-type-plot-sec) shows the distribution of encounters types.

```{r encntr-type-plot-sec, fig.cap="Distribution of Encounter Type."}
proportion_bar_plot(primary_enc,"ENCNTR_TYPE_CD","Distribution of Encounter Type.")
```

### Advanced Care Directive


```{r acd-tab-sec}
known_to_hf <- problem %>%
  dplyr::filter(stringr::str_detect(SOURCE_STRING,"(?i)Known to heart failure")) %>%
  dplyr::select(PERSON_ID) %>%
  dplyr::mutate(HAS_ALERT = TRUE)

kk<- primary_cohort %>%
  merge(.,known_to_hf , by="PERSON_ID",all=TRUE) %>%
  replace(is.na(.), FALSE)

prop_tab <- prop.table(table(kk$HAS_PRIMARY_HF,kk$HAS_ALERT))
prop_tab %>% 
  knitr::kable(caption = "Contigency table of Primary HF ICD10 and Advanced Care directive (ACD).",digits=2) %>%
  kableExtra::add_header_above(c("Has Primary\n ICD10"=1,"Has ACD"=2))%>%
  kable_styling(position = "center",full_width = FALSE)

```

Table \@ref(tab:acd-tab-sec) shows a contingency table of the number of patients with a primary ICD10 code who also have an Advanced Care Directive. **`r round(prop_tab[2,1]/sum(prop_tab[2,])*100,1)`**\% of patients with a primary HF ICD10 have an advanced care directive. Table xxx shows the cohort with either a primary or secondary HF diagnosis.

### Historical Diagnosis

Figure \@ref(fig:his-diag-plot-sec) shows whether or not, each person in the cohort has had a an encounter with a HF ICD10 diagnosis within the last three years. 

```{r his-diag-plot-sec, fig.cap="Distribution of historical diagnosis."}
his_hf_diag <- diagnosis_history %>%
  dplyr::filter(stringr::str_detect(SOURCE_IDENTIFIER,"(?i)I50") &
                  PERSON_ID %in% primary_cohort$PERSON_ID)

summary_his_diag <- extract_hf_diagnosis(his_hf_diag) %>%
  dplyr::right_join(primary_cohort,by="PERSON_ID") %>%
  dplyr::select(-HAS_PRIMARY_HF) %>%
  replace(is.na(.),FALSE) %>%
    dplyr::mutate(`No His. HF ICD10` = !`Congestive heart failure` & !`Heart failure, unspecified`
                  & !`Left ventricular failure`)
plot_logical_columns(summary_his_diag,"ICD10","Distribution of Historical ICD10")
```
### Orders

```{r}

echos <- orders %>%
  dplyr::filter(stringr::str_detect(CATALOG_CD,"(?i)cardio - echo") & ENCNTR_ID %in% primary_enc$ENCNTR_ID) %>%
  dplyr::group_by(ENCNTR_ID) %>%
  dplyr::arrange(desc(STATUS_DT_TM)) %>%
  dplyr::slice(1) %>%
  dplyr::select(ENCNTR_ID,CATALOG_CD,STATUS_DT_TM) %>%
  dplyr::right_join(
    dplyr::select(primary_enc,ENCNTR_ID,ARRIVE_DT_TM),by="ENCNTR_ID") %>%
  dplyr::mutate(HAS_ECHO = !is.na(CATALOG_CD),ECHO_DELTA = difftime(STATUS_DT_TM,ARRIVE_DT_TM,units = "days")) %>%
  dplyr::ungroup() %>%
  dplyr::select(-CATALOG_CD) %>%
  dplyr::rename(ECHO_START_DT_TM = STATUS_DT_TM)

```


```{r}
path_tests <- execute_query("SELECT * from PATHOLOGY_EVENT where CATALOG_CD = 'Brain Natriuretic Peptide' or CATALOG_CD = 'Iron Level'")

path_tests %<>%
  dplyr::filter(ENCNTR_ID %in% primary_enc$ENCNTR_ID) %>%
  dplyr::group_by(ENCNTR_ID,EVENT_CD) %>%
  dplyr::arrange(desc(EVENT_START_DT_TM)) %>%
  dplyr::slice(1) %>%
  dplyr::ungroup()



```

```{r order-prop-plot-sec, fig.cap="Number and proportion of Encounters with a relevant heart failure order, echocardiogram, brain natriuretic peptide test, serum iron test and transferrin saturation test."}
hf_order<-echos  %>%
  dplyr::left_join(
    extract_path_result(path_tests,"Brain Natriuretic peptide (BNP )", "BNP", c(-Inf,450,900,Inf),c("< 450","450-900","900+")), by="ENCNTR_ID"
  ) %>%
  dplyr::left_join(
    extract_path_result(path_tests,"Iron", "IRON",c(-Inf,9,30.4,Inf),c("< 9","9-30.4","30.4+")), by="ENCNTR_ID"
  ) %>%
  dplyr::left_join(
    extract_path_result(path_tests,"Transferrin Saturation", "TRANSFERRIN_SAT",c(-Inf,20,Inf),c("< 20","20+")), by="ENCNTR_ID"
  ) %>%
  dplyr::mutate(HAS_BNP = !is.na(BNP_EVENT_START_DT_TM),
                HAS_IRON = !is.na(IRON_EVENT_START_DT_TM),
                HAS_TRANSFERRIN_SAT = !is.na(TRANSFERRIN_SAT_EVENT_START_DT_TM))

plot_logical_columns(hf_order,"ORDERS","Proportion of Encounter with Order")


```


```{r echo-time-plot-sec,fig.cap="Distribution of the time taken for an echocardiogram."}

mean_delta <- mean(hf_order$ECHO_DELTA,na.rm = TRUE)
plot_histogram(hf_order %>% dplyr::filter(ECHO_DELTA < 75),
               "ECHO_DELTA","Distribution of time taken for Echo",
               mean_delta) + 
    xlab("Time take for echo since admission (days)")
```

Figure \@ref(fig:order-prop-plot) shows the proportion of encounters with a relevant heart failure order being echocardiogram, 
brain natriuretic peptide test, serum iron test and transferrin saturation test. Furthermore Figure \@ref(fig:echo-time-plot) shows the distribution of the time taken for an echocardiogram with a mean time of **`r round(mean_delta,1)`** days.


```{r path-test-plot-sec,fig.cap="Results of pathology tests."}


p1<-plot_pie_chart(hf_order,"BNP_BIN","BNP Results by Group")
p2<-plot_pie_chart(hf_order,"IRON_BIN","BNP Results by Group")
p3<-plot_pie_chart(hf_order,"TRANSFERRIN_SAT_BIN","BNP Results by Group")

#ggpubr::ggarrange(p1,p2,p3,nrow=3,ncol=1,labels=c("A","B","C"))


p1
p2
p3
```

### Forms


```{r form-plot-sec,fig.cap = "Proportion of Patients with an enrolment or referral form."}


form_primary_cohort <- primary_cohort %>%
  dplyr::left_join(
    extract_form(hf_form,'Heart Failure - Management of Cardiac Function Enrolment',"ENROLMENT"), by = c("PERSON_ID" = "ENROLMENT_PERSON_ID")
  ) %>%
  dplyr::left_join(
    extract_form(hf_form,'Heart Failure (MACARF) Referral Form',"REFERRAL"), by = c("PERSON_ID" = "REFERRAL_PERSON_ID")
  ) %>%
  tidyr::replace_na(list(ENROLMENT_HAS_FORM = FALSE, REFERRAL_HAS_FORM = FALSE))

plot_logical_columns(form_primary_cohort %>% dplyr::select(-HAS_PRIMARY_HF),"Form","Proportion with HF Form")
```
